---
title: "My Homework 1"
author: "Bin Liu"
date: "2023-09-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing excel file and adding header row

```{r}
library(readxl)
Typical_Employee_Survey_Data <- read_excel("C:/Users/charl/Desktop/Rutgers MBA Courses/Business Forecasting/Homeworks/Typical Employee Survey Data.xlsx", 
    col_names = c("Age","Gender","Job_Satisfaction","Job_Characteristics","Years_of_Employment","Promotion_in_Five_Years","Decision_Making_Involvement","Budgetary_Decision","Pride_in_Organization","Job_Loyalty","Workplace_Relationship"))
View(Typical_Employee_Survey_Data)

names(Typical_Employee_Survey_Data)
head(Typical_Employee_Survey_Data)
tail(Typical_Employee_Survey_Data)
```

# Summary and Boxplot

### Age

```{r}
summary(Typical_Employee_Survey_Data$Age)
fivenum(Typical_Employee_Survey_Data$Age)
boxplot(Typical_Employee_Survey_Data$Age)
```

### Gender

```{r}
summary(Typical_Employee_Survey_Data$Gender)
fivenum(Typical_Employee_Survey_Data$Gender)
boxplot(Typical_Employee_Survey_Data$Gender)
```

### Job_Satisfacation

```{r}
summary(Typical_Employee_Survey_Data$Job_Satisfaction)
fivenum(Typical_Employee_Survey_Data$Job_Satisfaction)
boxplot(Typical_Employee_Survey_Data$Job_Satisfaction)
```

### Job_Characteristics

```{r}
summary(Typical_Employee_Survey_Data$Job_Characteristics)
fivenum(Typical_Employee_Survey_Data$Job_Characteristics)
boxplot(Typical_Employee_Survey_Data$Job_Characteristics)
```

### Years_of_Employment

```{r}
summary(Typical_Employee_Survey_Data$Years_of_Employment)
fivenum(Typical_Employee_Survey_Data$Years_of_Employment)
boxplot(Typical_Employee_Survey_Data$Years_of_Employment)
```

### Promotion_in_Five_Years

```{r}
summary(Typical_Employee_Survey_Data$Promotion_in_Five_Years)
fivenum(Typical_Employee_Survey_Data$Promotion_in_Five_Years)
boxplot(Typical_Employee_Survey_Data$Promotion_in_Five_Years)
```

### Decision_Making_Involvement

```{r}
summary(Typical_Employee_Survey_Data$Decision_Making_Involvement)
fivenum(Typical_Employee_Survey_Data$Decision_Making_Involvement)
boxplot(Typical_Employee_Survey_Data$Decision_Making_Involvement)
```

### Budgetary_Decision

```{r}
summary(Typical_Employee_Survey_Data$Budgetary_Decision)
fivenum(Typical_Employee_Survey_Data$Budgetary_Decision)
boxplot(Typical_Employee_Survey_Data$Budgetary_Decision)
```

### Pride_in_Organization

```{r}
summary(Typical_Employee_Survey_Data$Pride_in_Organization)
fivenum(Typical_Employee_Survey_Data$Pride_in_Organization)
boxplot(Typical_Employee_Survey_Data$Pride_in_Organization)
```

### Job_Loyalty

```{r}
summary(Typical_Employee_Survey_Data$Job_Loyalty)
fivenum(Typical_Employee_Survey_Data$Job_Loyalty)
boxplot(Typical_Employee_Survey_Data$Job_Loyalty)
```

### Workplace_Relationship

```{r}
summary(Typical_Employee_Survey_Data$Workplace_Relationship)
fivenum(Typical_Employee_Survey_Data$Workplace_Relationship)
boxplot(Typical_Employee_Survey_Data$Workplace_Relationship)
```

# Converting the columns that need to be converted to Factors

### Converting Gender into a Factor

```{r}
Typical_Employee_Survey_Data$Gender <- as.factor(Typical_Employee_Survey_Data$Gender)
class(Typical_Employee_Survey_Data$Gender)
```

### Converting Job_Satisfaction into a Factor

```{r}
Typical_Employee_Survey_Data$Job_Satisfaction <- as.factor(Typical_Employee_Survey_Data$Job_Satisfaction)
class(Typical_Employee_Survey_Data$Job_Satisfaction)
```

### Converting Job_Characteristics into a Factor

```{r}
Typical_Employee_Survey_Data$Job_Characteristics <- as.factor(Typical_Employee_Survey_Data$Job_Characteristics)
class(Typical_Employee_Survey_Data$Job_Characteristics)
```

### Converting Promotion_in_Five_Years into a Factor

```{r}
Typical_Employee_Survey_Data$Promotion_in_Five_Years <- as.factor(Typical_Employee_Survey_Data$Promotion_in_Five_Years)
class(Typical_Employee_Survey_Data$Promotion_in_Five_Years)
```

### Converting Decision_Making_Involvement into a Factor

```{r}
Typical_Employee_Survey_Data$Decision_Making_Involvement <- as.factor(Typical_Employee_Survey_Data$Decision_Making_Involvement)
class(Typical_Employee_Survey_Data$Decision_Making_Involvement)
```

### Converting Budgetary_Decision into a Factor

```{r}
Typical_Employee_Survey_Data$Budgetary_Decision <- as.factor(Typical_Employee_Survey_Data$Budgetary_Decision)
class(Typical_Employee_Survey_Data$Budgetary_Decision)
```

### Converting Pride_in_Organization into a Factor

```{r}
Typical_Employee_Survey_Data$Pride_in_Organization <- as.factor(Typical_Employee_Survey_Data$Pride_in_Organization)
class(Typical_Employee_Survey_Data$Pride_in_Organization)
```

### Converting Job_Loyalty into a Factor

```{r}
Typical_Employee_Survey_Data$Job_Loyalty <- as.factor(Typical_Employee_Survey_Data$Job_Loyalty)
class(Typical_Employee_Survey_Data$Job_Loyalty)
```

### Converting Workplace_Relationship into a Factor

```{r}
Typical_Employee_Survey_Data$Workplace_Relationship <- as.factor(Typical_Employee_Survey_Data$Workplace_Relationship)
class(Typical_Employee_Survey_Data$Workplace_Relationship)
```

### Showing the columns changed to Factors

```{r}
head(Typical_Employee_Survey_Data)
```

# Plot the right visual for each column

### Age - Histogram

```{r}
hist(Typical_Employee_Survey_Data$Age, main = "Histogram of Age", xlab = "Age", ylab = "Number of Employee", col = "Skyblue")
```

### Gender - Bar chart

```{r}
barplot(table(Typical_Employee_Survey_Data$Gender), main = "Bar Chart of Gender", xlab = "Gender (1=Male, 2=Female)", ylab = "Number of Employee", col = "Plum")
```

### Job_Satisfaction - Bar Chart

```{r}
barplot(table(Typical_Employee_Survey_Data$Job_Satisfaction), main = "Bar Chart of Job Satisfaction", xlab = "Job Satisfaction (1=Very Satisfied, 4=Very Dissatisfied)", ylab = "Number of Employee", col = "Yellow")
```

### Job_Characteristics - Pie chart

```{r}
pie(table(Typical_Employee_Survey_Data$Job_Characteristics), main = "Pie Chart of Job Characteristics", labels = c("High Income","No Danger of Beinig Fired","Flexible Hours","Opportunities for Advancement","Enjoying the Work"))
```

### Years_of_Employment - Histogram

```{r}
hist(Typical_Employee_Survey_Data$Years_of_Employment, main = "Histogram of Years of Employment", xlab = "Years of Employment", ylab = "Number of Employee", col = "Green")
```

### Promotion_in_Five_Years - Pie chart

```{r}
pie(table(Typical_Employee_Survey_Data$Promotion_in_Five_Years), main = "Pie Chart of Promotion in Five Years", labels = c("Very Likely","Likely","Not Sure","Unlikely","Very Unlikely"))
```

### Decision_Making_Involvement - Bar chart

```{r}
barplot(table(Typical_Employee_Survey_Data$Decision_Making_Involvement), main = "Bar Chart of Decision Making Involvement", xlab = "Decision Making Involvement (1=Always, 5=Never)", ylab = "Number of Employee", col = "Aquamarine")
```

### Budgetary_Decision - Pie chart

```{r}
pie(table(Typical_Employee_Survey_Data$Budgetary_Decision), main = "Pie Chart of Budgetary Decision", labels = c("Yes","No"))
```

### Pride_in_Organization - Bar chart

```{r}
plot(Typical_Employee_Survey_Data$Pride_in_Organization, main = "Bar chart Pride in Organization", xlab = "Pride in Organization (1=Very Pround, 4=Not At All Proud)", ylab = "Number of Employee", col = "Lightsalmon")
```

### Job_Loyalty - Pie chart

```{r}
pie(table(Typical_Employee_Survey_Data$Job_Loyalty), main = "Pie Chart of Job Loyalty", labels = c("Very Likely","Likely","Not Sure","Unlikely","Very Unlikely"))
```

### Workplace_Relationship - Bar chart

```{r}
barplot(table(Typical_Employee_Survey_Data$Workplace_Relationship), main = "Bar Chart of Workplace Relationship", xlab = "Workplace Relationship (1=Very Good, 5=Very Bad)", ylab = "Number of Employee", col = "Mediumpurple")
```

# Scatter Plots for relationships

### Years of Employment vs Age

```{r}
plot(Typical_Employee_Survey_Data$Age, Typical_Employee_Survey_Data$Years_of_Employment, main = "Scatter Plot of Years of Employment vs Age", xlab = "Age", ylab = "Years of Employment", col = "Blue")
```

### Gender vs Age

```{r}
plot(Typical_Employee_Survey_Data$Age, Typical_Employee_Survey_Data$Gender, main = "Scatter Plot of Gender vs Age", xlab = "Age", ylab = "Gender (1=Male, 2=Female)", col = "Red")
```

### Job Satisfication vs Age

```{r}
plot(Typical_Employee_Survey_Data$Age, Typical_Employee_Survey_Data$Job_Satisfaction, main = "Scatter Plot of Job Satisfication vs Age", xlab = "Age", ylab = "Job Satisfication (1=Very Satisfied, 4=Very Dissatisfied)", col = "Purple")
```

### Promotion in Five Years vs Age

```{r}
plot(Typical_Employee_Survey_Data$Age, Typical_Employee_Survey_Data$Promotion_in_Five_Years, main = "Scatter Plot of Promotion in Five Years vs Age", xlab = "Age", ylab = "Promotion in Five Years (1=Very Likely, 5=Very Unlikely)", col = "Brown")
```

### Gender vs Years of Employment

```{r}
plot(Typical_Employee_Survey_Data$Years_of_Employment, Typical_Employee_Survey_Data$Gender, main = "Scatter Plot of Gender vs Years of Employmentr", xlab = "Years of Employment", ylab = "Gender (1=Male, 2=Female)", col = "Hotpink")
```

### Promotion in Five Years vs Years of Employment 

```{r}
plot(Typical_Employee_Survey_Data$Years_of_Employment, Typical_Employee_Survey_Data$Promotion_in_Five_Years, main = "Scatter Plot of Promotion in Five Years vs Years of Employment", xlab = "Years of Employment", ylab = "Promotion in Five Years (1=Very Likely, 5=Very Unlikely)", col = "Orange")
```

### Job Loyalty vs Years of Employment 

```{r}
plot(Typical_Employee_Survey_Data$Years_of_Employment, Typical_Employee_Survey_Data$Job_Loyalty, main = "Scatter Plot of Job Loyalty vs Years of Employment", xlab = "Years of Employment", ylab = "Job Loyalty (1=Very likely, 5=Very Unlikely)", col = "Darkviolet")
```

### Job Loyalty vs Pride in Organization

```{r}
plot(as.numeric(Typical_Employee_Survey_Data$Pride_in_Organization), as.numeric (Typical_Employee_Survey_Data$Job_Loyalty), main = "Scatter Plot of Job Loyalty vs Pride in Organization", xlab = "Pride in Organization (1=Very Proud, 5=Not At All Pround)", ylab = "Job Loyalty (1=Very likely, 5=Very Unlikely)", col = "Olivedrab")
```

### Job Loyalty vs Workplace Relationship

```{r}
plot(as.numeric(Typical_Employee_Survey_Data$Workplace_Relationship), as.numeric (Typical_Employee_Survey_Data$Job_Loyalty), main = "Scatter Plot of Job Loyalty vs Workplace Relationship", xlab = "Workplace Relationship (1=Very Good, 5=Very Bad)", ylab = "Job Loyalty (1=Very Likely, 5= Very Unlikely)", col = "Magenta4")
```

### Number of Employee who are proud of the organization vs Number of Employe who are satisfied with their Job

```{r}
plot(as.numeric(Typical_Employee_Survey_Data$Pride_in_Organization), main="Number of employee who are pround vs satisfied with thier job", col="Red", xlab="Number of Employee", ylab="Job Satisfaction (Blue), Pride in Organization (Red)")
points(as.numeric(Typical_Employee_Survey_Data$Job_Satisfaction), col="Blue")
```