---
title: "My Homework 5"
author: "Bin Liu"
date: "2023-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing data for Electric Vehicle Popultion Size History

```{r}
library(readr)
Electric_Vehicle_Population_Size_History <- read_csv("~/Documents/Fall 2023 - Business Forecasting/Datasets/Electric_Vehicle_Population_Size_History.csv")
View(Electric_Vehicle_Population_Size_History)
``` 

# Creating a time series for Electric Vehicle (EV) Total

```{r}
library(fpp)
library(fpp2)
library(TTR)

Electric_Vehicle_Population_Total_ts <- ts(Electric_Vehicle_Population_Size_History$`Electric Vehicle (EV) Total`, frequency = 12, start = c(2017,1))
Electric_Vehicle_Population_Total_ts

attributes (Electric_Vehicle_Population_Total_ts)
plot(Electric_Vehicle_Population_Total_ts)
Acf(Electric_Vehicle_Population_Total_ts)
```

# Forecasting Models

## Average Model

#### The Average Model takes the average of all historical points. The plot shows that the avearge of the total registration of electric vehicles in the State of Washington is 66,273.58. This model will take this average to forecast for the next 24 months, which is the blue straight line in the plot. The shaded area shows the confidence interval that the future values may lie between this area. As we can see from the plot, this is not a good forecasting model for this data based on the trend since the forecasted values are way below the current trend.

```{r}
mean_forecast <- meanf(Electric_Vehicle_Population_Total_ts, 24)
plot(mean_forecast)
Acf(mean_forecast$residuals)
attributes(mean_forecast)
summary(mean_forecast)
```

## Naive Model

#### The Naive Model takes the last month's data to forecast for the next 24 months as shown in the blue straight line in the plot. The confidence interval shows the future values may lie within this shaded area based on this prediction. This forecasting model does not appear to be a good model based on the trend. 

```{r}
naive_forecast <- naive(Electric_Vehicle_Population_Total_ts, 24)
plot(naive_forecast)
Acf(naive_forecast$residuals)
summary(naive_forecast)
```

## Random Walk Forecast (RWF)

#### Similar to the Naive Model, the Random Walk Forecast Model also takes the last month's data to forecast for the values in the next 24 months with the same values for the confidence interval. Again, it is not considered a good forecasting model based on the trend. 

```{r}
rwf_forecast <- rwf(Electric_Vehicle_Population_Total_ts,24)
plot(rwf_forecast)
Acf(rwf_forecast$residuals)
summary(rwf_forecast)
```

## A drift with RWF

#### The Random Walk Forecast Model with a drift shows an upward trend for the next 24 months. The confidence interval also has an increased values as opposed to the Random Walk Forecast Model. This shows a better forecast based on the trend of the time series.

```{r}
rwf_forecast_drift <- rwf(Electric_Vehicle_Population_Total_ts,24, drift=TRUE)
plot(rwf_forecast_drift)
Acf(rwf_forecast_drift$residuals)
summary(rwf_forecast_drift)
```

## Seasonal Naive

#### The time series shows an upward trend and does not show any seasonality trend. Therefore, the Seasnoal Naive Model is not a good model. Based on the plot, it appears that it takes the upward trend from the recent months then a downward trend, then a similar upward trend again to forecast the values for the next 24 months. This model does not represent the increasing upward trend of the time series.

```{r}
snaive_forecast <- snaive(Electric_Vehicle_Population_Total_ts,24)
plot(snaive_forecast)
Acf(snaive_forecast$residuals)
summary(snaive_forecast)
```

## Moving Averages

### Moving average with order of 3

#### Based on the ACF, it shows that the recent data are more significant than the historical data. Therefore, the lower order moving average model provides a better forecast than higher order moving average model since the larger weight is given to the more recent period.

```{r}
MA3_forecast <- ma(Electric_Vehicle_Population_Total_ts,order=3)
plot(MA3_forecast)
```

### Moving average with order of 6

#### The moving average model with order of 6 still provides a good forecast for the upward trend although it smooths out the data more than the previous model with order of 3. It gives more weight to more historical data which appears to be less significant. This is not as good as the previous model with the lower order. 

```{r}
MA6_forecast <- ma(Electric_Vehicle_Population_Total_ts,order=6)
plot(MA6_forecast)
```

### Moving average with order of 12

#### The larger order of the moving average model, the smoother effect on the plot. As we can see from the plot, it gives more weight to more historical data as comparing to the moving average model with order of 3 and 6. Therefore, it is not as good as the previous two models with the lower order. 

```{r}
MA12_forecast <- ma(Electric_Vehicle_Population_Total_ts,order=12)
plot(MA12_forecast)
```

### Moving average with order of 18

#### As mentioned above, the larger order of moving average puts more weight on more historical data. Therefore, the moving average with large order does not represent well for this time series.

```{r}
MA18_forecast <- ma(Electric_Vehicle_Population_Total_ts,order=18)
plot(MA18_forecast)
```

## Holtwinters

#### The Holtwinters Model without any smoothing constants provides a pretty good fit to the original data plot. The residuls are very small based on the comparison of the actual and forecasted data (black line vs red line).

```{r}
HW_forecast <- HoltWinters(Electric_Vehicle_Population_Total_ts)
plot(HW_forecast)
attributes(HW_forecast)
summary(HW_forecast)
plot(HW_forecast)
```

#### The Holtwinters Model with exponential smoothing for trend and non-seasonal model provides a decent fit. However, the residual are greater than the previous model with no smoothing constants. 


```{r}
SSE_Simple <- HoltWinters(Electric_Vehicle_Population_Total_ts,beta=FALSE,gamma=FALSE)
attributes(SSE_Simple)
plot(SSE_Simple)
SSE_Simple$SSE
head(SSE_Simple$fitted)
```
#### The Holtwinters Model with non-exponential smoothing for trend and non-seasonal model provides a pretty nice fit and the residuals are small among the acutal and forecasted data.

```{r}
SSE_Simple_1 <- HoltWinters(Electric_Vehicle_Population_Total_ts,beta=TRUE,gamma=FALSE)
attributes(SSE_Simple_1)
plot(SSE_Simple_1)
SSE_Simple_1$SSE
head(SSE_Simple_1$fitted)
```
#### The Holtwinters Model with non-exponential smoothing for trend and seasonal model also provides a pretty nice fit and the residuals are small among the acutal and forecasted data.

```{r}
SSE_Simple_2 <- HoltWinters(Electric_Vehicle_Population_Total_ts,beta=TRUE,gamma=TRUE)
attributes(SSE_Simple_2)
plot(SSE_Simple_2)
SSE_Simple_2$SSE
head(SSE_Simple_2$fitted)
```

#### The Holtwinters Model with exponential smoothing for trend and seasonal model provides quite similar forecast with the exponential smoothing for trend and non-seaonal model. In both models, the residuals are greater than the other Holtwinters models displayed above. 

```{r}
SSE_Simple_3 <- HoltWinters(Electric_Vehicle_Population_Total_ts,beta=FALSE,gamma=TRUE)
attributes(SSE_Simple_3)
plot(SSE_Simple_3)
SSE_Simple_3$SSE
head(SSE_Simple_2$fitted)
```

# Plot all in a single chart

```{r}
plot(mean_forecast, col="black")
lines(naive_forecast$mean,col="red")
lines(rwf_forecast$mean, col="green")
lines(rwf_forecast_drift$mean, col="hotpink")
lines(snaive_forecast$mean, col="blue")
lines(MA6_forecast, col="purple")
lines(MA12_forecast, col="orange")
lines(MA18_forecast, col="yellow")
```

# Decomposition

```{r}
ets_forecast <- ets(Electric_Vehicle_Population_Total_ts)
plot(ets_forecast)
attributes(ets_forecast)
ets_forecast$mse
```

# Accuracy measures for different forecasting models

#### I pick the Mean Absolute Percentage Error (MAPE) as the accuracy measure for this time series. By comparing the MAPE with different models listed below, the Moving Average with order of 3 shows the least MAPE. This indicates that the Moving Average with order of 3 is the best model based on the accuracy comparison.

```{r}
library(forecast)
```

## Average Model

```{r}
accuracy(mean_forecast)
```

## Naive Model

```{r}
accuracy(naive_forecast)
```

## Random Walk Forecast Model

```{r}
accuracy(rwf_forecast)
```

## Random Walk Forecast with a Drift Model

```{r}
accuracy(rwf_forecast_drift)
```

## Seasonal Naive Model

```{r}
accuracy(snaive_forecast)
```

## Moving Averages with different orders (3, 6, 12 and 18)

```{r}
accuracy(MA3_forecast,Electric_Vehicle_Population_Total_ts, h=24)
accuracy(MA6_forecast,Electric_Vehicle_Population_Total_ts, h=24)
accuracy(MA12_forecast,Electric_Vehicle_Population_Total_ts, h=24)
accuracy(MA18_forecast,Electric_Vehicle_Population_Total_ts, h=24)
```

## Holtwinters with different combinations of smoothing trends and seasonal models

```{r}
forecast_ets_hw <- forecast(HW_forecast, h=24)
accuracy(forecast_ets_hw)

forecast_ets_SSE_Simple <- forecast(SSE_Simple, h=24)
accuracy(forecast_ets_SSE_Simple)

forecast_ets_SSE_Simple_1 <- forecast(SSE_Simple_1, h=24)
accuracy(forecast_ets_SSE_Simple_1)

forecast_ets_SSE_Simple_2 <- forecast(SSE_Simple_2, h=24)
accuracy(forecast_ets_SSE_Simple_2)

forecast_ets_SSE_Simple_3 <- forecast(SSE_Simple_3, h=24)
accuracy(forecast_ets_SSE_Simple_3)
```

# Seasonal Decomposition of Time Series by Loess

```{r}
stl_decomp <- stl(Electric_Vehicle_Population_Total_ts, s.window = "periodic")
plot(stl_decomp)
attributes (stl_decomp)
stl_decomp$weights
stl_decomp$call
stl_decomp$win
stl_decomp$deg
stl_decomp$jump
stl_decomp$inner
stl_decomp$outer
```

# Seasonal adjustment 

```{r}
tmp <- seasadj(stl_decomp)
plot(Electric_Vehicle_Population_Total_ts, col="black")
lines(tmp, col="red")
```

# Forecast of Seasonal Decomposition of Time Series by Loess

```{r}
f_stl <- forecast(stl_decomp)
f_stl <- forecast(stl_decomp, h=24)
plot(f_stl)
accuracy(f_stl)
attributes(f_stl)
Acf(f_stl$residuals)
```

# Classical Seasonal Decomposition by Moving Averages

```{r}
decomp_ev <- decompose(Electric_Vehicle_Population_Total_ts)
plot(decomp_ev)
attributes (decomp_ev)
```

# Classical Seasonal Decomposition by Moving Averages - Multiplicative

```{r}
decomp_evm <- decompose(Electric_Vehicle_Population_Total_ts, type="multiplicative")
plot(decomp_evm)
attributes (decomp_evm)
```

# Forecast of Classical Seasonal Decomposition by Moving Averages

### By using two different seasonal decomposition methods, the forecasting accuracy measure for Classical Seasonal Decomposition by Moving Averages has lower MAPE. The ACF plot for residuals also looks better for Classical Seasonal Decomposition by Moving Averages as the lines are all within the blue dash lines on the plot.

```{r}
f_decomp <- forecast(seasadj(decomp_ev), h=24)
plot(f_decomp)
accuracy(f_decomp)
Acf(f_decomp$residuals)
```
